Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-09-01 supplement.tex
started admixture discussion

Low coverage (4-8x) Illumina short read data were generated for the Nama,
Gumuz, Amhara and Oromo populations as part of the African Diversity Reference
Panel (Sanger / Wellcome Trust) \citep{Gurdasani2015-qy,Pagani2015-pz} and
approved through a secondary data analysis agreement for this project. Briefly,
raw reads were aligned to GRCH37 with BWAmem, duplicates
marked with Picard MarkDuplicates, reads were realigned around indels with GATK
RealignerTargetCreater/IndelRealigner followed by BQSR with dbSNP 137.
Contamination checks were performed requiring that FREEMIX $<0.05$;
contamination checks resulted in the elimination of 22 Nama samples. We note
that the high heterozygosity in these genomes both due to inherent genetic
diversity and admixture may have violated base assumptions for this
heterozygosity check. Genomes were then variant called with GATK3.2 Unified
Genotyper \citep{DePristo2011-up} using joint calling across 2,478 individuals within
the African Diversity Reference Panel (ADRP) dataset with a
minimum base quality of 17. Data were merged with 1000
Genomes Phase 3 \citep{1000_Genomes_Project_Consortium2015-zq} using the union
of sites identified with bcftools isec (-n+1) \citep{Danecek2021-kc} then refined
with Variant Recalibrator with a truth sensitivity threshold of 99.5\%. HapMap
III and dbSNP 138 served as known sites while 1000 Genomes Phase 1 Omni2.5 and
Phase 1 genomic SNP served as the training set. After VR, no batch effects were
observed along PC1 and PC2 for 1000 Genomes vs. ADRP. Phasing on the combined
dataset was performed via SHAPEIT2 \citep{Delaneau2013-aw} and utilized the duoHMM
option for duos and trios. We present new whole-genome data for 82 Nama individuals 
available under accession number EGAD00001006198.
Second- and third-degrees relatives were inferred from reconstructed
pedigrees with Omni2.5 SNP array data.
Among these 82 Nama samples, we down-sampled the individuals to minimize
close relatives and admixture, such that 44 Nama genomes were retained.
Individuals with $>70\%$ estimated Khoe-San ancestry were retained for
analysis, after partitioning ancestry into $k=6$ clusters with \texttt{ADMIXTURE}
\citep{Alexander2009-sw} where alternative ancestries represent European, West
African, Near Eastern, and East African gene flow.

>>>>

Low coverage (4-8x) Illumina short read data were generated for the Nama,
Gumuz, Amhara and Oromo populations as part of the African Diversity Reference
Panel (Sanger / Wellcome Trust) \citep{Gurdasani2015-qy,Pagani2015-pz} and
approved through a secondary data analysis agreement for this project. Briefly,
raw reads were aligned to GRCH37 with BWAmem, duplicates
marked with Picard MarkDuplicates, reads were realigned around indels with GATK
RealignerTargetCreater/IndelRealigner followed by BQSR with dbSNP 137.
Contamination checks were performed requiring that FREEMIX $<0.05$;
contamination checks resulted in the elimination of 22 Nama samples. We note
that the high heterozygosity in these genomes both due to inherent genetic
diversity and admixture may have violated base assumptions for this
heterozygosity check. Genomes were then variant called with GATK3.2 Unified
Genotyper \citep{DePristo2011-up} using joint calling across 2,478 individuals within
the African Diversity Reference Panel (ADRP) dataset with a
minimum base quality of 17. Data were merged with 1000
Genomes Phase 3 \citep{1000_Genomes_Project_Consortium2015-zq} using the union
of sites identified with bcftools isec (-n+1) \citep{Danecek2021-kc} then refined
with Variant Recalibrator with a truth sensitivity threshold of 99.5\%. HapMap
III and dbSNP 138 served as known sites while 1000 Genomes Phase 1 Omni2.5 and
Phase 1 genomic SNP served as the training set. After VR, no batch effects were
observed along PC1 and PC2 for 1000 Genomes vs. ADRP. Phasing on the combined
dataset was performed via SHAPEIT2 \citep{Delaneau2013-aw} and utilized the duoHMM
option for duos and trios. We present new whole-genome data for 82 Nama individuals 
available under accession number EGAD00001006198.
Second- and third-degrees relatives were inferred from reconstructed
pedigrees with Omni2.5 SNP array data.
Among these 82 Nama samples, we down-sampled the individuals to minimize
close relatives and admixture, such that 44 Nama genomes were retained.
Individuals with $>70\%$ estimated Khoe-San ancestry were retained for
analysis, after partitioning ancestry into $k=6$ clusters with an initial \texttt{ADMIXTURE}
\citep{Alexander2009-sw} analysis on \sgcomment{TODO: Brenna complete}, where alternative ancestries represent European, West
African, Near Eastern, and East African gene flow. 
Other \texttt{ADMIXTURE} analyses
presented in this manuscript were restricted to a subset of 1157 individuals \sgcomment{TODO:  Aaron complete}.
